Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2022-03-11 ms/data_readme.Rmd
Update data readme
- automatically generate txt
- automatically zip figshare files

---
output: rtf_document
editor_options:
  chunk_output_type: console
---

---
bibliography: 
- `r here::here("ms/references.yaml")`
---

@@

```{r setup, include = FALSE, cache = FALSE}
# TO CONVERT TO PLAIN TEXT:
# after rendering to rtf,
# 1. Copy the contents of the rtf file, paste to txt.
# 2. Replace all single line breaks with double line breaks
# (in Atom search-replace with grep: \n for \n\n)
# 3. Hard-wrap all text at 80 characters (in Atom with autoflow: cmd-alt-q)
# (optional: add name after first line)
# 4. Save as "README.txt"

knitr::opts_chunk$set(
  echo = FALSE, message = FALSE,
  warning = FALSE, results = "hide", cache = FALSE)

# Load objects from targets workflow (all paths to data files)
tar_load(
  c(
    "ms_functions",
    "ebihara_2019_zip_file",
    "ftol_zip_file",
    "ja_climate_data_file",
    "japan_map_file",
    "deer_range_file",
    "japan_ferns_comm_full_file",
    "inext_res_file",
    "occ_point_data_summary_file",
    "plastome_calibration_dates",
    "redundancy_by_res_file",
    "japan_ferns_shape_full_file",
    "raw_trait_data_file",
    "japan_map_points_file",
    "protected_areas_file",
    "calibration_dates_file",
    "japan_ferns_biodiv_figshare_file",
    "comm_ferns_figshare_file",
    "shape_ferns_figshare_file",
    "fern_traits_figshare_file",
    "japan_fern_phylogram_figshare_file",
    "japan_fern_tree_figshare_file",
    "japan_fern_tree_uncollapsed_figshare_file",
    "refs_yaml"
    )
  )
```

@@

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\- GENERAL INFORMATION \-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

@@

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

@@

Citation for and links to publications that cite or use the data: Nitta JH,
Mishler BD, Iwasaki W, Ebihara A (2022) Spatial phylogenetics of Japanese ferns:
Patterns, processes, and implications for conservation FIXME: add journal when published

@@

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

@@

Phylogenetic analysis was conducted with maximum likelihood in IQ-TREE v1.6.12
[@Nguyen2015] by combining plastid *rbcL* sequences of each taxon with a
globally sampled data matrix (Nitta et al, in prep). Next, dating analysis was
carried out using treePL v1.0 [@Smith2012] with `r n_fossil_points` fossil
calibration points after @Testo2016a. The dated phylogeny was then trimmed to
include Japanese taxa only.

@@

For full methods, see Nitta JH, Mishler BD, Iwasaki W, Ebihara A (2022) Spatial
phylogenetics of Japanese ferns: Patterns, processes, and implications for
conservation FIXME: Add journal when published

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

@@

```{r results = "asis"}
  deer_range %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "range", "Source of range data. '1978' = 1978 survey, '2003' = 2003 survey, 'estimated' = range estimated from model based on 2003 survey data",
      "geom", "Vector describing shape and position of range polygon"
  ))
```

@@

```{r ja-climate-data-cols, results = "asis"}
  ja_climate_data %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "temp", "Mean annual temperature, in units of 10 * degrees celsius (BIO1)",
      "temp_season", "Temperature seasonality (standard deviation ×100) (BIO4)",
      "precip", "Annual precipitation (mm) (BIO12)",
      "precip_season", "Precipitation seasonlaity (coefficient of variation) (BIO15)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

@@

```{r japan-ferns-inext-results-cols, results = "asis"}
  inext_res %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "m", "Sample size (number of individuals)",
      "method", "Method used to estimate diveristy",
      "order", "Diversity order of Hill number (here 0, equivalent to species richness)",
      "qD", "Estimated diversity",
      "qD.LCL", "Lower 95% confidence limit on estimated diversity",
      "qD.UCL", "Upper 95% confidence limit on estimated diversity",
      "SC", "Estimated sampling completeness",
      "SC.LCL", "Lower 95% confidence limit on sampling completeness",
      "SC.UCL", "Upper 95% confidence limit on sampling completeness")
  )
```

@@

`r fs::path_file(occ_point_data_summary_file)`: Summary of raw specimen data for
native, non-hybrid ferns of Japan. Raw specimen data consists of one row per
specimen, with species name, latitude, longitude, and collection date (raw
specimen data not included here to protect endangered species). Raw data were
filtered to remove duplicate specimens and specimens that occurred outside of
the second-degree mesh (approximately 10 km sq grid-cell) map of Japan
(http://www.biodic.go.jp/trialSystem/EN/shpddl.html)

@@

Numeric trait data formatted as a series of numbers separated by colons (':'),
unless missing (0 or 3). For example, 1:9.4:12:14:16. The starting '1:' does not
mean anything. The next series of numbers correspond to the outside minimum,
normal minimum, normal maximum and outside maximum. For example, a plant leaf
may normally be in the range of 4-5 cm, but sometimes as low as 2 cm or as high
as 6 cm. This would be written 2:4:5:6.

@@

`r fs::path_file(japan_ferns_biodiv_figshare_file)` (contained in "results.zip"):
Biodiversity statistics of native, non-hybrid ferns and environmental variables
in Japan. Biodiversity metrics calculated as described in Nitta et al. 2022. 
Climatic (temperature and preciptation) variables
calculated as described for japan_climate.gpkg. Includes one row with missing
environmental data and one outlier for % apomixis that were removed prior to
spatial modeling analysis in Nitta et al. 2022.

@@

```{r japan-ferns-biodiv-figshare-cols, results = "asis"}
  japan_ferns_biodiv_figshare %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "lat", "Latitude of grid-cell centroid",
      "long", "Longitude of grid-cell centroid",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "fd_obs_z", "Z-score (standard effect size) of observed functional diversity (FD)",
      "pd_obs_z", "Z-score of observed phylogenetic diversity (PD)",
      "pe_obs_z", "Z-score of observed phylogenetic endemisim (PE)",
      "rfd_obs_z", "Z-score of observed relative FD",
      "rpd_obs_z", "Z-score of observed relative PD",
      "richness_obs_p_upper", "Richness percentile",
      "fd_obs_p_upper", "FD percentile",
      "pd_obs_p_upper", "PD percentile",
      "pe_obs_p_upper", "PE percentile",
      "pd_signif", "Significance of PD compared to null distribution, two-sided test at alpha 0.05",
      "rpd_signif", "Significance of RPD compared to null distribution, two-sided test at alpha 0.05",
      "fd_signif", "Significance of FD compared to null distribution, two-sided test at alpha 0.05",
      "rfd_signif", "Significance of RFD compared to null distribution, two-sided test at alpha 0.05",
      "pe_signif", "Significance of PE compared to null distribution, two-sided test at alpha 0.05",
      "taxonomic_cluster", "Membership in cluster based on taxonomic distances",
      "phylo_cluster", "Membership in cluster based on phylogenetic distances",
      "endem_type", "Endemism type determined by CANAPE",
      "lat_area", "Area of 1-degree latitudinal band around grid-cell",
      "temp", "Annual mean temperature, in units of degrees celsius * 10",
      "temp_season", "Temperature seasonality",
      "precip", "Annual precipitation, measured in mm",
      "precip_season", "Precipitation seasonality",
      "percent_apo", "Percent apomictic taxa"
    ))
```

@@

\-\-\-

@@

Generate this README file.

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

References

>>>>

---
output: md_document
editor_options:
  chunk_output_type: console
---

---
bibliography: 
- `r here::here("ms/data_readme_refs.yaml")`
---

@@

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(
  echo = FALSE, message = FALSE,
  warning = FALSE, results = "hide", cache = FALSE)

# Load objects from targets workflow (all paths to data files)
tar_load(
  c(
    "ms_functions",
    "ebihara_2019_zip_file",
    "ftol_zip_file",
    "ja_climate_data_file",
    "japan_map_file",
    "deer_range_file",
    "japan_ferns_comm_full_file",
    "inext_res_file",
    "occ_point_data_summary_file",
    "plastome_calibration_dates",
    "redundancy_by_res_file",
    "japan_ferns_shape_full_file",
    "raw_trait_data_file",
    "japan_map_points_file",
    "protected_areas_file",
    "calibration_dates_file",
    "japan_ferns_biodiv_figshare_file",
    "comm_ferns_figshare_file",
    "shape_ferns_figshare_file",
    "fern_traits_figshare_file",
    "japan_fern_phylogram_figshare_file",
    "japan_fern_tree_figshare_file",
    "japan_fern_tree_uncollapsed_figshare_file",
    "refs_data_readme_yaml"
    )
  )
```

@@

------------------- GENERAL INFORMATION -----------------

@@

--------------------------

@@

Citation for and links to publications that cite or use the data: Nitta JH,
Mishler BD, Iwasaki W, Ebihara A (2022) Spatial phylogenetics of Japanese ferns:
Patterns, processes, and implications for conservation.
American Journal of Botany. FIXME: add DOI, page numbers when available

@@

--------------------

@@

Phylogenetic analysis was conducted with maximum likelihood in IQ-TREE v1.6.12
[@Nguyen2015] by combining plastid rbcL sequences of each taxon with a
globally sampled data matrix (FTOL v0.0.1; Nitta et al., in prep.; 
https://github.com/fernphy/ftol/). Next, dating analysis was
carried out using treePL v1.0 [@Smith2012] with `r n_fossil_points` fossil
calibration points after @Testo2016a. The dated phylogeny was then trimmed to
include Japanese taxa only.

@@

For full methods, see Nitta JH, Mishler BD, Iwasaki W, Ebihara A (2022) Spatial
phylogenetics of Japanese ferns: Patterns, processes, and implications for
conservation. American Journal of Botany. 
FIXME: add DOI, page numbers when available

--------------------------

@@

```{r deer-range-cols, results = "asis"}
  deer_range %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "range", "Source of range data. '1978' = 1978 survey, '2003' = 2003 survey, 'estimated' = range estimated from model based on 2003 survey data",
      "geom", "Vector describing shape and position of range polygon"
  ))
```

@@

```{r ja-climate-data-cols, results = "asis"}
  ja_climate_data %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "temp", "Mean annual temperature, in units of 10 * degrees celsius (BIO1)",
      "temp_season", "Temperature seasonality (standard deviation ×100) (BIO4)",
      "precip", "Annual precipitation (mm) (BIO12)",
      "precip_season", "Precipitation seasonality (coefficient of variation) (BIO15)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

@@

```{r japan-ferns-inext-results-cols, results = "asis"}
  inext_res %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "m", "Sample size (number of individuals)",
      "method", "Method used to estimate diversity",
      "order", "Diversity order of Hill number (here 0, equivalent to species richness)",
      "qD", "Estimated diversity",
      "qD.LCL", "Lower 95% confidence limit on estimated diversity",
      "qD.UCL", "Upper 95% confidence limit on estimated diversity",
      "SC", "Estimated sampling completeness",
      "SC.LCL", "Lower 95% confidence limit on sampling completeness",
      "SC.UCL", "Upper 95% confidence limit on sampling completeness")
  )
```

@@

`r fs::path_file(occ_point_data_summary_file)`: Summary of raw specimen data for
native, non-hybrid ferns of Japan. Raw specimen data consists of one row per
specimen, with species name, latitude, longitude, and collection date (raw
specimen data not included here). Raw data were
filtered to remove duplicate specimens and specimens that occurred outside of
the second-degree mesh (approximately 10 km sq grid-cell) map of Japan
(http://www.biodic.go.jp/trialSystem/EN/shpddl.html)

@@

Numeric trait data formatted as a series of numbers separated by colons (':'),
unless missing (0 or 3). For example, 1:9.4:12:14:16. The starting '1:' does not
mean anything. The next series of numbers correspond to the outside minimum,
normal minimum, normal maximum and outside maximum. For example, a plant leaf
may normally be in the range of 4-5 cm, but sometimes as low as 2 cm or as high
as 6 cm. This would be written 1:2:4:5:6.

@@

`r fs::path_file(japan_ferns_biodiv_figshare_file)` (contained in "results.zip"):
Biodiversity statistics of native, non-hybrid ferns and environmental variables
in Japan. Biodiversity metrics calculated as described in Nitta et al. 2022. 
Climatic (temperature and preciptation) variables
calculated as described for japan_climate.gpkg. Includes three rows with missing 
data and one outlier for % apomixis that were removed prior to
spatial modeling analysis.

@@

```{r japan-ferns-biodiv-figshare-cols, results = "asis"}
  japan_ferns_biodiv_figshare %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "lat", "Latitude of grid-cell centroid",
      "long", "Longitude of grid-cell centroid",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "fd_obs_z", "Z-score (standard effect size) of observed functional diversity (FD)",
      "pd_obs_z", "Z-score of observed phylogenetic diversity (PD)",
      "pe_obs_z", "Z-score of observed phylogenetic endemism (PE)",
      "rfd_obs_z", "Z-score of observed relative FD",
      "rpd_obs_z", "Z-score of observed relative PD",
      "richness_obs_p_upper", "Richness percentile",
      "fd_obs_p_upper", "FD percentile",
      "pd_obs_p_upper", "PD percentile",
      "pe_obs_p_upper", "PE percentile",
      "pd_signif", "Significance of PD compared to null distribution, two-sided test at alpha 0.05",
      "rpd_signif", "Significance of RPD compared to null distribution, two-sided test at alpha 0.05",
      "fd_signif", "Significance of FD compared to null distribution, two-sided test at alpha 0.05",
      "rfd_signif", "Significance of RFD compared to null distribution, two-sided test at alpha 0.05",
      "pe_signif", "Significance of PE compared to null distribution, two-sided test at alpha 0.05",
      "taxonomic_cluster", "Membership in cluster based on taxonomic distances",
      "phylo_cluster", "Membership in cluster based on phylogenetic distances",
      "endem_type", "Endemism type determined by CANAPE",
      "lat_area", "Area of 1-degree latitudinal band around grid-cell",
      "temp", "Annual mean temperature, in units of degrees celsius * 10",
      "temp_season", "Temperature seasonality",
      "precip", "Annual precipitation, measured in mm",
      "precip_season", "Precipitation seasonality",
      "percent_apo", "Percent apomictic taxa"
    ))
```

@@

2022-03-11

- Use Mollweide projection for grid-cells
- Don't include targets cache (move to https://github.com/joelnitta/japan_ferns_spatial_phy_cache)
- Fix spelling errors in README text
- Fix missing '1:' in LUCID trait data example
- Fix description of rows with missing data in japan_ferns_biodiv.csv
- Automate generation of plain text file

@@

- Generate this README file.

--------------------------

REFERENCES
